A comparative analysis of penalized linear mixed models in structured genetic data

### 2020-12-03 main.tex
Section 4 meeting notes

For each $\bX$ matrix, $\bZ$ was a corresponding incidence matrix of 1s and 0s identifying subpopulation, or founding population membership, which assigns the corresponding environmental effect found in $\bgamma$. Various $\bgamma$ patterns were considered in order to investigate the effects of distinct and complex patterns of environmental confounding. Firstly, we considered two different patterns of effect magnitudes, which we refer to as exponential and dichotomous. We use the term dichotomous to describe two effect sizes each experienced by half of the population, regardless of the number of subpopulations, while exponential effects refer to a unique effect size for each subpopulation with exponentially increasing magnitudes. Secondly, we considered whether individuals in genetically similar subpopulations would experience similar environmental effects. We refer to environmental effects that are similar among genetically similar subpopulations as concordant, and to those that are not as discordant.

@@

In both scenarios, LMM-lasso is more robust to allele frequency differences than PC-lasso due to the presence of \anna{cryptic relatedness in addition to population structure in the Empirical data.} \anna{max absolute kinship? histograms of off-diag kinship?} This echoes the superior TPRs of LMM-lasso compared to that of the PC-lasso in the case of the Empirical data in Table \ref{tab:var_sel}.

>>>>

For each $\bX$ matrix, $\bZ$ was a corresponding incidence matrix of 1s and 0s identifying subpopulation, or founding population membership, which assigns the corresponding environmental effect found in $\bgamma$. Various $\bgamma$ patterns were considered in order to investigate the effects of distinct and complex patterns of environmental confounding. Firstly, we considered two different patterns of effect magnitudes, which we refer to as exponential and dichotomous. We use the term dichotomous to describe two effect sizes each experienced by half of the population, regardless of the number of subpopulations, while exponential effects refer to a unique effect size for each subpopulation with exponentially increasing magnitudes. Secondly, we considered whether individuals in genetically similar subpopulations would experience similar environmental effects. We refer to environmental effects that are similar among genetically similar subpopulations as concordant, and to those that are not as discordant. \anna{We examined these but found they didn't make a huge difference in performance so we don't focus on them.}

@@

\anna{LMM-lasso seems to exhibit even stronger control of this than PC-lasso.}

%In both scenarios, LMM-lasso is more robust to allele frequency differences than PC-lasso due to the presence of \anna{finer grained relatedness in addition to population stratification in the Empirical data.} This echoes the superior TPRs of LMM-lasso compared to that of the PC-lasso in the case of the Empirical data in Table \ref{tab:var_sel}.
